---
title: "Life Expectancy at Birth App"
author: "Pablo Rosales"
date: "12/20/2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Life Expectancy at Birth

- Life expectancy at birth is defined as "... a statistical measure of the average time an organism is expected to live, based on the year of its birth, its current age and other demographic factors including gender."  (Source:  Wikipedia)

- Worldwide, life expectancy at birth shows an incremental trend for most countries.

- The trends vary by country and by sex.

## The App

- The "Life Expectancy at Birth App" allows the user to explore this variable interactively.

- Posible analyses include: 
  - Life expectancy at birth for a given country, sex and year of birth.
  - Life expectancy trend in a given country.
  - Life expectancy in a given country compared to World statistics. 

- It is powered by data provided by the Institute for Health Metrics and Evaluation, University of Washington (https://www.gapminder.org/data/documentation/gd004/)

## Demo:  Life expectancy at birth in Guatemala for a person born in 1978

```{r echo = FALSE, include = FALSE}
  # ----------------------------------- Load Libraries ----------------------------------- #
    library(sqldf)
    library(dplyr)
    library(reshape2)
    library(ggplot2)
    
   # ----------------------------------- Load Data Sets ----------------------------------- #
  raw_life_expectancy_at_birth_male_df <- read.csv("/Users/prosales/Documents/Personales/09 - Developing Data Products/final_project/data/Life_expectancy_at_birth_MALE.csv")
  
  # ----------------------------------- Males Set Preparation ----------------------------------- #
  # rename colums
  raw_life_expectancy_at_birth_male_col_names <- colnames(raw_life_expectancy_at_birth_male_df) 
  std_life_expectancy_at_birth_male_col_names <- gsub("X", "", raw_life_expectancy_at_birth_male_col_names)
  colnames(raw_life_expectancy_at_birth_male_df) <- std_life_expectancy_at_birth_male_col_names
  
  # standardise countries names
  std_life_expectancy_at_birth_male_df <- raw_life_expectancy_at_birth_male_df %>% mutate (country = toupper(country))
  
  # convert column year to row
  std_2_life_expectancy_at_birth_male_df <- melt(std_life_expectancy_at_birth_male_df, id = (c(("country"))) )
  colnames(std_2_life_expectancy_at_birth_male_df) <- c("country", "year", "life_expectancy_at_birth")
  
  std_2_life_expectancy_at_birth_male_df$year <- as.integer(as.character(std_2_life_expectancy_at_birth_male_df$year))
  
  std_2_life_expectancy_at_birth_male_df$sex = "MALE"

  # ----------------------------------- All  Set Preparation ----------------------------------- #
  all_life_expectancy_at_birth_df <- std_2_life_expectancy_at_birth_male_df
  
  # ----------------------------------- Getting User's Choices ----------------------------------- #
  sex_choice <- "MALE"
  country_choice <- "GUATEMALA"
  year_of_birth_choice <- 1978
  
  # ----------------------------------- Filtering  User's Choice ----------------------------------- #f
  resulting_row_by_choice <- all_life_expectancy_at_birth_df[
  (
    (all_life_expectancy_at_birth_df$country == country_choice) &
    (all_life_expectancy_at_birth_df$year    == year_of_birth_choice) & 
    (all_life_expectancy_at_birth_df$sex     == sex_choice)
  ), 
  ]
  
  life_expectancies_by_country_df <- all_life_expectancy_at_birth_df [
    (
      (all_life_expectancy_at_birth_df$country == country_choice) &
        (all_life_expectancy_at_birth_df$sex   == sex_choice)
    ), 
  ]
  
  
  
  
  # ----------------------------------- Life Expectancy at Birth in Years for Country ----------------------------------- #
  life_expectancies_by_country_qplot <- qplot(
    x = year, 
    y = life_expectancy_at_birth,
    data = life_expectancies_by_country_df,
    main = paste("Period Life Expectancy at Birth by Year in ", country_choice),
    xlab = "Year",
    ylab = "Life Expectancy at Birth in Years"
  )
  
  life_expectancies_by_country_qplot <- life_expectancies_by_country_qplot + theme(axis.text.x = element_text(angle = 90, hjust = 1))
  
  life_expectancies_by_country_qplot <- life_expectancies_by_country_qplot + geom_vline(xintercept = 1978, linetype="dashed", color = "red")
  
  
```

``` {r echo = TRUE}
life_expectancies_by_country_qplot
```

## Conclusion and Invitation

- Understanding World through data can be an enlightening endeavour. 

- This app represents a tool to understand a fundamental variable to understand human progress, you are invited to use it !


